### 2020-12-20 paper/sex_selection_spacing-ver5.tex
Update Table titles and footnotes and clean up comments
Discuss statistical significance in birth spacing differences

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant 
changes in India.
Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the 25th, 50th, and 75th percentile birth interval durations in 
months, the sex ratio, and the probability of having a birth for each spell
by education levels and area of residence.%
\footnote{
Results for urban women without education, rural women with 12 or more years of 
education, and the fourth spell for women with 12 or more years of education are
in the online appendix because of relatively small samples.
}
The sex ratio graphs also show the natural sex ratio, approximately 51.2\% boys
\citep{ben-porath76b,jacobsen99,Portner2015b}.
The underlying values with bootstrapped standard errors are available in the online 
appendix.

@@

To separate the effects of the introduction of sex selection and the other changes in
India, I first discuss how the lengths of birth intervals have changed in situations where, 
based on sex ratios, there appears to be less sex selection.
The group broadly covers women with no education, regardless of their children's sex 
composition, and women with any education who have one or two sons already. 
Despite the apparent lower level of sex selection, son preference is still evident with 
the shortest spacing when they have only girls. 
Notably, for those who appear the least likely to use sex selection---rural women with no 
education---the difference in birth interval length across sex compositions has grown over 
time as spacing when sons are present has become longer.

@@

A clear illustration of how the combination of son preference and the introduction of 
sex selection affected birth spacing comes from the third spell of the best-educated urban 
women.
With two girls, almost 80\% of the third births are boys, and the 75th percentile 
birth interval length is close to 70 months. 
This interval is about 13 months longer than if they had at least one son and represents 
an increase of almost 21 months over the four decades.

@@

Not surprisingly, given these changes, the third spell for the best-educated women shows
the clearest reversal in the spacing pattern; the birth intervals with two girls are
consistently longer than the intervals with one or two boys, no matter the percentile used.
A similar reversal, although more muted, occurred for the third spell for urban women 
with 1--7 years of education and both urban and rural women with 8--11 years of 
education.

@@

Sex selection, however, appears to be behind the most substantial increases in birth spacing. 
The best-educated women with two girls had the most biased sex ratio and the most 
significant increase in birth interval lengths. 
Over the four decades, the median birth interval length for this group increased by almost 
15 months, 
and the 75th percentile birth interval length increased by a staggering 21 months, most of 
that within a decade.

@@

% Questions
The main question I address in this paper is how birth spacing responded to the 
significant changes in India between 1972 and 2016, particularly the spread of sex 
selection.
I also examine two related questions.
First, did the changes in birth spacing bias the standard fertility estimates for India? 
Second, what is the relationship between infant mortality and the changes in birth spacing 
and sex selection?

% What has happened to spacing and why
The most substantial lengthening of birth intervals came from the best-educated women,
likely because of substantial use of sex selection combined with falling fertility.
Take, for example, women with 12 or more years of education who had two girls. 
As the sex ratio reached almost 80\% boys, the expected median birth interval length
increased by almost 15 months, and the 75th percentile interval length increased by 21 months.
Most of the increase in the long intervals came immediately after the introduction of sex 
selection in India.

Some of the lengthening in birth intervals are so large that we even observed a reversal of
the traditional spacing pattern for some groups: 
For these groups, when there are no sons, we now see the longest, rather than the shortest, 
birth intervals, consistent with substantial use of sex selection.
The women who are the least likely to use sex selection, as indicated by their sex ratios
at birth, still show the traditional spacing pattern with short spacing in the absence of
sons.

@@

% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 25th Percentile Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:p25_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 25th Percentile Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:p25_rural}
% \end{figure}
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 50th Percentile Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:p50_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 50th Percentile Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:p50_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in Average Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:avg_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in Average Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:avg_rural}
% \end{figure}

@@

% \section{Survival Curves Conditional on Parity Progression for All Education and Spell Groups}
% 
% \setcounter{figure}{0}
% \setcounter{table}{0}
% 
% 
% % [FIRST SPELL]
% 
% % \subsection{First Spell}
% % 
% % \input{../figures/appendix_spell1_low.tex}
% % 
% % \input{../figures/appendix_spell1_med.tex}
% % 
% % \input{../figures/appendix_spell1_high.tex}
% % 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{No Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_low_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_low_rural_pps}} \\
% % \caption*{1-7 Years of Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_med_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_med_rural_pps}} \\
% % \caption*{8 or more Years of Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_highest_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_highest_rural_pps}} 
% % \caption{Survival curves conditional on progression to first birth; start point is month of marriage}
% % \label{fig:results_spell1_pps}
% % \end{figure}
% % 
% % 
% % \clearpage
% % \newpage
% % 
% \subsection{Second Spell}
% 
% 
% % \input{../figures/appendix_spell2_low.tex}
% 
% 
% % PARITY PROGRESSION SURVIVAL CURVES - Low
% 
% % Low education
% \vspace*{-0.75cm}
% \begin{figure}[hp!]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2005]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with no education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_low_pps}
% \end{figure}
% 
% 
% 
% 
% % SPELL 2 - URBAN - MEDIUM
% 
% % \input{../figures/appendix_spell2_med.tex}
% 
% % PARITY PROGRESSION SURVIVAL - MEDIUM
% 
% 
% % Low education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with 1-7 years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_med_pps}
% \end{figure}
% 
% 
% 
% % High education
% 
% % \input{../figures/appendix_spell2_high.tex}
% 
% % PARITY PROGRESSION SURVIVAL - High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with eight to eleven years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_high_pps}
% \end{figure}
% 
% 
% % Highest education
% 
% % PARITY PROGRESSION SURVIVAL - Highest education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }

>>>>

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant 
changes in India.
Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the 25th, 50th, and 75th percentile birth interval durations in 
months, the sex ratio, and the probability of having a birth for each spell
by education levels and area of residence.%
\footnote{
Results for urban women without education, rural women with 12 or more years of 
education, and the fourth spell for women with 12 or more years of education are
in the online appendix because of relatively small samples.
}
The sex ratio graphs also show the natural sex ratio, approximately 51.2\% boys
\citep{ben-porath76b,jacobsen99,Portner2015b}.
The underlying values with bootstrapped standard errors and tests for statistically
significant differences across sex compositions are available in the online appendix.

@@

To separate the effects of the introduction of sex selection and the other changes in
India, I first discuss how the lengths of birth intervals have changed in situations where, 
based on sex ratios, there appears to be less sex selection.
The group broadly covers women with no education, regardless of their children's sex 
composition, and women with any education who have one or two sons already.

Despite the apparent lower level of sex selection, son preference is still evident with 
the shortest spacing when they have only girls. 
Notably, for those who appear the least likely to use sex selection---rural women with no
education---birth spacing is almost uniformly statistically significantly longer when at
least one son is present than if no son is present. 
Furthermore, the difference in birth interval length across sex compositions has grown over
time as spacing when sons are present has become longer.

@@

A clear illustration of how the combination of son preference and the introduction of 
sex selection affected birth spacing comes from the third spell of the best-educated urban 
women.
With two girls, almost 80\% of the third births are boys, and the 75th percentile 
birth interval length is close to 70 months, an increase of almost 21 months over the four
decades. 
This interval has gone from being statistically significantly shorter than the intervals 
with at least one son to now being a statistically significant 13 months longer.

@@

Not surprisingly, given these changes, the third spell for the best-educated women shows
the clearest reversal in the spacing pattern; the birth intervals with two girls are
consistently statistically significantly longer than the intervals with one or two boys, 
no matter the percentile used.
A similar reversal, although more muted, occurred for the third spell for urban women 
with 1--7 years of education and both urban and rural women with 8--11 years of 
education.

@@

However, sex selection appears to be behind the most substantial increases in birth spacing. 
The best-educated women with two girls had the most biased sex ratio and the most 
significant increase in birth interval lengths. 
Over the four decades, the median birth interval length for this group increased by almost
15 months. 
Even more striking, the 75th percentile birth interval length increased by a staggering 21
months, most of that within a decade of the introduction of sex selection.

@@

% Questions
The main question I address in this paper is how birth spacing responded to the 
significant changes in India between 1972 and 2016, particularly the spread of sex 
selection.
I also examine two related questions.
First, did changes in birth spacing make us overestimate the decline in cohort fertility?
Second, what is the relationship between infant mortality and the changes in birth spacing 
and sex selection?

% What has happened to spacing and why
The most substantial lengthening of birth intervals came from the best-educated women,
likely because of substantial use of sex selection combined with falling fertility.
Take, for example, women with 12 or more years of education who had two girls. 
As the sex ratio reached almost 80\% boys, the expected median birth interval length
increased by almost 15 months, and the 75th percentile interval length increased by 21 months.
Most of the lengthening in the long intervals came immediately after the introduction of sex 
selection in India.

Some of the lengthening in birth intervals are so large that we even observed a reversal of
the traditional spacing pattern for some groups: 
For these groups, when there are no sons, we now see statistically significant longer,
rather than statistically significant shorter, birth intervals than if there is at least
one son.
The women who are the least likely to use sex selection, as indicated by their sex ratios
at birth, still show the traditional spacing pattern with short spacing in the absence of
sons.

@@

@@
